\section*{References}

\begin{enumerate}
\item Gepner R, Mihovilovic Skanata M, Berber NM, Dacosta M, Kadaba AD, Rubin GM, Clark DA (2015). Computations underlying \textit{Drosophila} photo-taxis, odor-taxis, and multi-sensory integration. \textit{eLife} 4:e06229.

\item Gershow M, Berck M, Mathew D, Luo L, Kane EA, Carlson JR, Samuel ADT (2012). Controlling airborne cues to study small animal navigation. \textit{Nature Methods} 9:290--296.

\item Gomez-Marin A, Stephens GJ, Louis M (2011). Active sampling and decision making in \textit{Drosophila} chemotaxis. \textit{Nature Communications} 2:441.

\item Hernandez-Nunez L, Belina J, Klein M, Si G, Claus L, Carlson JR, Samuel ADT (2015). Reverse-correlation analysis of navigation dynamics in \textit{Drosophila} larva using optogenetics. \textit{eLife} 4:e06225.

\item Kane EA, Gershow M, Afonso B, Larderet I, Klein M, Carter AR, de Bivort BL, Sprecher SG, Samuel ADT (2013). Sensorimotor structure of \textit{Drosophila} larva phototaxis. \textit{PNAS} 110:E3868--E3877.

\item Klein M, Afonso B, Vonner AJ, Hernandez-Nunez L, Berck M, Taber CJ, Sprecher SG, Gershow M, Garrity PA, Samuel ADT (2015). Sensory determinants of behavioral dynamics in \textit{Drosophila} thermotaxis. \textit{PNAS} 112:E220--E229.

\item Paninski L (2004). Maximum likelihood estimation of cascade point-process neural encoding models. \textit{Network: Computation in Neural Systems} 15:243--262.

\item Pillow JW, Shlens J, Paninski L, Sher A, Litke AM, Chichilnisky EJ, Simoncelli EP (2008). Spatio-temporal correlations and visual signalling in a complete neuronal population. \textit{Nature} 454:995--999.
\end{enumerate}

