---
title: "Video Quality Analysis Report"
subtitle: "Flaco Owl Documentary Footage - Fifth Avenue Studios"
author: ""
date: today
format:
  pdf:
    documentclass: article
    geometry: margin=1in
    fontsize: 11pt
    colorlinks: true
    keep-tex: true
    toc: false
    number-sections: true
---

# Executive Summary

This report provides a comprehensive technical and narrative analysis of video footage documenting Flaco the owl's first-night landing on Fifth Avenue after being freed from the Central Park Zoo. The footage has been evaluated for documentary production suitability, content significance, and licensing valuation.

**Key Findings:**

| Finding | Value |
|---------|-------|
| Total Duration | 15.83 seconds |
| Usable Footage | 15.83 seconds (100%) |
| High Quality Segments | 6.93 seconds (43.8%) |
| Narrative Significance | 100/100 (CRITICAL) |
| Estimated Value Range | \$2,226 - \$2,713 |

# Technical Specifications

| Property | Value |
|----------|-------|
| Resolution | 720 × 1280 pixels (Full HD 1080p) |
| Frame Rate | 30.00 fps |
| Duration | 15.83 seconds |
| Total Frames | 475 frames |
| File Size | 4.76 MB |
| Codec | Unknown |
| Color Depth | 8-bit |

# Visual Quality Metrics

```{r}
#| label: fig-quality-radar
#| fig-cap: "Video Quality Metrics Radar Chart - Normalized scores showing overall quality across key dimensions"
#| fig-width: 8
#| fig-height: 8
knitr::include_graphics("figures/quality_metrics_radar.pdf")
```

## Core Quality Metrics

| Metric | Score | Assessment |
|--------|-------|------------|
| Sharpness | 33.65 | Very Soft/Blurry |
| Brightness | 72.8/255 | Underexposed |
| Contrast | 44.53 | High Contrast |
| Noise Level | 5.32 | Very Clean |
| Exposure | 98.0/100 | Excellent Exposure |
| Stability | 5.43 | Unstable/Shaky |
| Smoothness | 0.937 | Very Smooth |
| Color Compression | 0.876 | Minimal Compression |

## Quality Assessment

**Strengths:**

| Strength | Description |
|----------|-------------|
| Resolution | High resolution (Full HD 1080p) suitable for broadcast |
| Exposure | Well-exposed footage with minimal over/under exposure |
| Noise Level | Clean footage with minimal noise (5.32) |
| Smoothness | Excellent temporal smoothness (0.937) |
| Compression | Minimal color compression artifacts (0.876) |

**Issues Requiring Post-Processing:**

| Issue | Description | Required Action |
|-------|-------------|-----------------|
| Sharpness | Footage is soft/blurry (sharpness: 33.65) | Requires sharpening |
| Stability | Camera shake present (stability: 5.43) | Requires stabilization |
| Brightness | Slightly underexposed (brightness: 72.8/255) | May need color correction |

# Content Significance Analysis

## Subject Detection

```{r}
#| label: fig-content-timeline
#| fig-cap: "Content Detection Timeline - When the subject (bird/owl) appears throughout the video duration"
#| fig-width: 14
#| fig-height: 4
knitr::include_graphics("figures/content_timeline.pdf")
```

| Metric | Value |
|--------|-------|
| Frames Analyzed | 32 frames |
| Frames with Subject (Bird/Owl) | 14 frames (43.8%) |
| Frames with Dog | 0 frames |
| Frames with Interaction | 0 frames |
| Subject Visible Throughout | No |
| Interaction Moments Detected | No |
| Rare/Unique Event | Yes |

## Event Significance

This footage documents a **rare and historic event**: Flaco the owl's first landing on Fifth Avenue on the night he was freed from the Central Park Zoo. This represents a critical moment in the narrative arc of the documentary, capturing the first public appearance of Flaco after his escape.

# Narrative Significance (Story Value)

## Narrative Score: 100/100 (CRITICAL)

```{r}
#| label: fig-narrative
#| fig-cap: "Narrative Significance Score Components - Breakdown of factors contributing to the critical story importance score"
#| fig-width: 10
#| fig-height: 8
knitr::include_graphics("figures/narrative_significance.pdf")
```

| Factor | Contribution |
|--------|--------------|
| Rare/Unique Event | First-night landing on Fifth Ave |
| Historic Moment | First appearance after zoo escape |
| Subject Visibility | Subject visible in 44% of footage |
| Multiple-Use Potential | Can be used multiple times in doc |
| Story Importance | CRITICAL |

**Multiple-Use Potential:** This footage can be strategically reused throughout the documentary in:

| Use Case | Description |
|----------|-------------|
| Opening Sequence | Can be used in documentary opening |
| Transitional Moments | Suitable for scene transitions |
| Key Narrative Scenes | Can feature in critical story moments |
| Closing Sequences | Appropriate for documentary conclusion |

# Usability Analysis

```{r}
#| label: fig-usability
#| fig-cap: "Usability Breakdown - Total duration, usable seconds, and high-quality segments"
#| fig-width: 10
#| fig-height: 6
knitr::include_graphics("figures/usability_breakdown.pdf")
```

## Per-Second Breakdown

| Category | Seconds | Percentage |
|----------|---------|------------|
| Total Duration | 15.83 | 100.0% |
| Usable Seconds | 15.83 | 100.0% |
| High Quality Seconds | 6.93 | 43.8% |

**Usable Segments:** 1 continuous segment (0.00s - 15.83s)

## Clear, Focused Segments

No continuous clear/focused segments of 1 second or longer were identified. The footage requires post-processing (sharpening and stabilization) to achieve optimal clarity for broadcast use.

# Valuation Guidance

```{r}
#| label: fig-valuation-breakdown
#| fig-cap: "Valuation Rate Calculation - Step-by-step breakdown showing how multipliers affect the final rate per second"
#| fig-width: 12
#| fig-height: 8
knitr::include_graphics("figures/valuation_breakdown.pdf")
```

```{r}
#| label: fig-value-range
#| fig-cap: "License Fee Value Range - Conservative estimate, agreed fee, and optimistic estimate comparison"
#| fig-width: 10
#| fig-height: 6
knitr::include_graphics("figures/value_range.pdf")
```

## Rate Calculation

| Factor | Multiplier | Value |
|--------|------------|-------|
| Base Rate (per usable second) | 1.00× | \$50.00 |
| Quality Multiplier | 0.50× | |
| Resolution Multiplier | 1.00× | |
| Content Significance Multiplier | 1.50× | |
| Narrative Significance Multiplier | 2.50× | |
| Multiple-Use Multiplier | 1.50× | |
| **Final Rate per Second** | | **\$140.62** |

## Estimated Value Range

| Estimate Type | Calculation | Value |
|---------------|-------------|-------|
| Conservative | All usable seconds × \$140.62 | \$2,226 |
| Optimistic | Premium segments at 1.5× rate | \$2,713 |
| **Recommended Range** | | **\$2,226 - \$2,713** |

## Valuation Notes

| Note | Description |
|------|-------------|
| Usability | High usability (100% usable) - most of the clip is production-ready |
| Archival Value | Rare/unique event adds significant archival and narrative value |
| Post-Processing | Lower technical quality may require post-processing, factor into negotiations |
| Story Value | Critical story moment with high reuse potential increases value |
| Licensing Premium | Multiple-use licensing should command premium rate |

# Recommendations

## For Licensing Negotiations

| Recommendation | Description |
|----------------|-------------|
| Base Value | Start with conservative estimate (\$2,226) as floor |
| Premium Value | Use optimistic estimate (\$2,713) for exclusive/premium use |
| Multiple Uses | Factor in 1.5× multiplier for multiple-use licensing |
| Post-Processing | Consider that footage requires stabilization and sharpening |
| Narrative Value | Emphasize critical story importance and archival significance |

## For Production Use

| Aspect | Recommendation |
|--------|----------------|
| Suitability | Footage is suitable for documentary use with post-processing |
| Stabilization | Recommend stabilization software (e.g., Adobe Warp Stabilizer) |
| Sharpening | Recommend sharpening filters to improve clarity |
| Color Correction | Color correction may enhance underexposed areas |
| Usage Flexibility | Footage can be used multiple times throughout documentary structure |

# Conclusion

This footage represents a critical moment in Flaco the owl's story, documenting his first landing on Fifth Avenue after being freed from the Central Park Zoo. While the technical quality requires post-processing, the narrative significance, rarity, and multiple-use potential make this footage highly valuable for documentary production.

The estimated licensing value range of \$2,226 - \$2,713 reflects both the technical quality considerations and the substantial narrative value of this historic moment.

\newpage

# Appendix: Score Computation Methodology

## Overall Quality Score

```{r}
#| label: fig-quality-components
#| fig-cap: "Quality Score Components - Weighted contribution of each metric to the overall quality score"
#| fig-width: 10
#| fig-height: 6
knitr::include_graphics("figures/quality_score_components.pdf")
```

The overall quality score is calculated using a weighted average of normalized metrics:

$$Q = 0.25 \cdot S_n + 0.20 \cdot E_n + 0.15 \cdot C_n + 0.15 \cdot N_n + 0.25 \cdot St_n$$

Where:

$$S_n = \min\left(100, \frac{\bar{S}}{500} \times 100\right) \quad \text{(Sharpness, normalized)}$$

$$E_n = \bar{E} \quad \text{(Exposure score, 0-100)}$$

$$C_n = \min\left(100, \frac{\bar{C}}{40} \times 100\right) \quad \text{(Contrast, normalized)}$$

$$N_n = \max\left(0, 100 - \frac{\bar{N}}{50} \times 100\right) \quad \text{(Noise, normalized)}$$

$$St_n = \max\left(0, 100 - \frac{\bar{St}}{5} \times 100\right) \quad \text{(Stability, normalized)}$$

## Sharpness Calculation

Sharpness is measured using Laplacian variance:

$$S = \text{Var}(\nabla^2 I)$$

Where $\nabla^2 I$ is the Laplacian of the grayscale image $I$.

## Brightness Assessment

Average brightness is calculated as:

$$B = \frac{1}{N} \sum_{i,j} I_{gray}(i,j)$$

Assessment thresholds:

| Category | Threshold |
|---------|----------|
| Well-exposed | $100 \leq B \leq 180$ |
| Underexposed | $50 \leq B < 100$ |
| Very dark | $B < 50$ |
| Overexposed | $B > 200$ |

## Contrast Calculation

RMS contrast is calculated as:

$$C = \sqrt{\frac{1}{N} \sum_{i,j} (I_{gray}(i,j) - \mu)^2}$$

Where $\mu$ is the mean brightness.

## Noise Level Estimation

Noise level is estimated using the standard deviation of the Laplacian:

$$N = \sigma(\nabla^2 I)$$

## Stability Calculation

Stability is measured using optical flow magnitude:

$$St = \frac{1}{M} \sum_{i,j} \sqrt{u(i,j)^2 + v(i,j)^2}$$

Where $(u, v)$ are the optical flow vectors between consecutive frames.

## Smoothness Calculation

Smoothness combines histogram similarity and color consistency:

$$Sm = \frac{H_{corr} + C_{cons}}{2}$$

Where:

$$H_{corr} = \text{Histogram correlation between frames}$$

$$C_{cons} = 1 - \frac{\bar{\Delta H}}{180} \quad \text{(Color consistency)}$$

## Color Compression Score

Compression score is calculated as:

$$CC = 1 - \frac{B_h + B_s + B_v}{3}$$

Where $B_h$, $B_s$, $B_v$ are banding scores for hue, saturation, and value channels respectively.

## Narrative Significance Score

Narrative score is calculated as:

$$NS = \min(100, \sum_{i} w_i \cdot f_i)$$

Where $w_i$ are weights and $f_i$ are factors:

| Factor | Points |
|--------|--------|
| Rare/unique event | +40 points |
| Historic moment | +30 points |
| Subject visibility (>30%) | +15 points |
| Clear focused segments (≥3s) | +10 points |
| Multiple clear segments | +5 points |
| Interaction moments | +10 points |
| Multiple-use potential | +20 points |

## Valuation Rate Calculation

The final rate per usable second is calculated as:

$$R = R_{base} \times M_q \times M_r \times M_c \times M_n \times M_m$$

Where:

$$R_{base} = \$50 \quad \text{(Base rate)}$$

$$M_q = \text{Quality multiplier (0.5-2.0)}$$

$$M_r = \text{Resolution multiplier (0.5-1.5)}$$

$$M_c = \text{Content significance multiplier (1.0-1.8)}$$

$$M_n = \text{Narrative significance multiplier (1.0-2.5)}$$

$$M_m = \text{Multiple-use multiplier (1.0-1.5)}$$

## Total Value Calculation

$$V_{conservative} = R \times T_{usable}$$

$$V_{optimistic} = R \times 1.5 \times T_{high} + R \times (T_{usable} - T_{high})$$

Where:

| Variable | Description |
|----------|------------|
| $T_{usable}$ | Total usable seconds |
| $T_{high}$ | High quality seconds |

## Technical Details

### Analysis Parameters

| Parameter | Value |
|-----------|-------|
| Frames sampled | 32 frames (2 frames per second) |
| Analysis method | Frame-by-frame quality assessment |
| Object detection | YOLOv8n model |
| Minimum segment duration | 1.0 second |
| Usability threshold | 35/100 points |
| High quality threshold | 60/100 points |

### Software and Tools

| Tool | Purpose |
|------|---------|
| OpenCV 4.12.0 | Video processing |
| Ultralytics YOLOv8 | Object detection |
| NumPy | Numerical computations |
| Custom algorithms | Quality metrics analysis |

